---
title: "SIndexRCFS Workflow Integration Examples"
author: "SIndexRCFS Contributors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIndexRCFS Workflow Integration Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SIndexRCFS)
```

## Overview

This vignette shows two practical integration patterns:

1. Permanent Sample Plot (PSP) site productivity estimation.
2. Treelist preparation for growth and yield model inputs.

The examples use generic sample datasets included in the package under `inst/examples`.

## 1) PSP Workflow Example

Typical PSP programs maintain dominant height and age observations by plot and measurement year.
Using those fields, SIndexRCFS can derive site index and years to breast height.

```{r psp-load}
psp_path <- system.file("examples", "psp_sample.csv", package = "SIndexRCFS")
psp <- read.csv(psp_path, stringsAsFactors = FALSE)
psp
```

```{r psp-calc}
psp$site_index_m <- mapply(
  function(sp, age, age_type, h, est_type) {
    ht_age_to_si(age = age, age_type = age_type, height = h, si_est_type = est_type, species = sp)
  },
  psp$species, psp$bh_age, psp$age_type, psp$dom_height_m, psp$si_est_type
)

psp$y2bh_years <- mapply(
  function(sp, si) si_to_y2bh(site_index = si, species = sp),
  psp$species, psp$site_index_m
)

psp
```

```{r psp-summary}
aggregate(
  psp[, c("site_index_m", "y2bh_years")],
  by = list(plot_id = psp$plot_id),
  FUN = mean
)
```

### Why This Helps

- **Site index (`site_index_m`)** gives a comparable productivity metric among plots.
- **Years to breast height (`y2bh_years`)** supports age conversions and growth model setup.

## 2) Treelist Workflow for Growth and Yield Inputs

A common workflow is to estimate stand-level productivity from dominant age/height,
then attach those attributes to each tree record used by a growth model.

```{r treelist-load}
stands_path <- system.file("examples", "stand_inputs_sample.csv", package = "SIndexRCFS")
trees_path <- system.file("examples", "treelist_sample.csv", package = "SIndexRCFS")

stands <- read.csv(stands_path, stringsAsFactors = FALSE)
trees <- read.csv(trees_path, stringsAsFactors = FALSE)

stands
trees
```

```{r treelist-calc}
stands$site_index_m <- mapply(
  function(sp, age, h) ht_age_to_si(age = age, age_type = 1, height = h, species = sp),
  stands$dom_species, stands$dom_bh_age, stands$dom_height_m
)

stands$y2bh_years <- mapply(
  function(sp, si) si_to_y2bh(site_index = si, species = sp),
  stands$dom_species, stands$site_index_m
)

stands$proj_height_m <- mapply(
  function(sp, age, si, y2bh) si_age_to_ht(iage = age, age_type = 1, site_index = si, y2bh = y2bh, species = sp),
  stands$dom_species, stands$projection_age, stands$site_index_m, stands$y2bh_years
)

stands
```

```{r treelist-join}
model_treelist <- merge(
  trees,
  stands[, c("stand_id", "site_index_m", "y2bh_years", "projection_age", "proj_height_m")],
  by = "stand_id",
  all.x = TRUE,
  sort = FALSE
)

model_treelist
```

### Why This Helps

- The treelist receives consistent stand productivity fields.
- Model inputs can include both current tree attributes and site productivity descriptors.

## Run As Scripts

You can run the equivalent script versions directly:

```r
source(system.file("examples", "psp_workflow_example.R", package = "SIndexRCFS"))
source(system.file("examples", "treelist_workflow_example.R", package = "SIndexRCFS"))
```

